---
title: "Project 4 - Algorithm Implementation and Evaluation"
author: "Guo, Yajie  Li, Yiran  Shao, Nina  Wu, Chenyun  Xuan, Sijian"
date: "11/27/2017"
output: html_document
---

## Step 0: Load the packages, specify directories

```{r}

```

## Step 1: Load, process and clean the data

```{r}

```

## Step 2: Implement memory-based algorithms

1) Similarity Weight

Spearman Correlation
```{r}

```

Entropy
```{r}

```

Mean-square-difference
```{r}

```

SimRank
```{r}

```

Significance Weighting
```{r}

```

Variance Weighting
```{r}

```

Selecting Neighbors
```{r}

```

Rating Normalization

Once the neighborhood has been selected, the ratings from those neighbors are combined to compute a prediction, after possibly scaling the ratings to a common distribution. An extension to the GroupLens algorithm is to account for the differences in spread between users' rating distributions by converting ratings to z-scores. And computing a weighted average of the z-scores as following.
```{r}

```

Obviously, we can get predictions in this step. In next step, we will compare the actual value in test data to prediction. 
## Step 3: Implement model-based algorithms

```{r}

```

## Step 4: Evaluation

Ranked scoring for the first data set.
```{r}

```

Mean absolute error(MAE) for the second data set.

Mean absolute error is the average of absolute value of the difference between the prediction and actual value from test data set. The calculation results are as follows.
```{r}

```

we can easily find that the algorithm which possesses the lowest MAE is the most efficient collaborative filtering algorithm. Here, the best one is     .

ROC sensitivity for the second data set.
```{r}

```

## Conclusion

