---
title: "Significance Weighting"
author: "Sijian Xuan"
date: "November 24, 2017"
output: html_document
---
##### read data
###movietrain_a = read.csv("../output/movie_train_a.csv")
####movietrain length should be 12773985
#### This is for MSweb data
```{r}
webtrain_a = read.csv("../output/webtrain_a.csv", header = FALSE)
webtrain_a = unlist(webtrain_a)
sqrt(length(webtrain_a)) # = 4151
```
### This is for movie data
```{r}
movietrain_a = read.csv("../output/movie_train_a.csv", header = FALSE)
movietrain_a = unlist(movietrain_a)
movie_train_a = matrix(0, 5055,5055)
movie_train_a[upper.tri(movie_train_a, diag=FALSE)] = unlist(movietrain_a)
movie_train_a = t(movie_train_a)
movie_train_a[upper.tri(movie_train_a, diag=FALSE)] = unlist(movietrain_a)
# diagonal is 0
```
### Note: the values in the webtrain_a and in movietrain_a are number of co-reated items pairwise.
##### choose the number of co-rated items
```{r}
print(mean(webtrain_a))
print(sd(webtrain_a))
print(mean(movie_train_a))
print(sd(movie_train_a))
```
## Web dataset
### It is reasonable that we choose 5, 8, 9 and 10 as the number of co-rated items.
### n = 5
```{r}
webtrain_a_n_5 = webtrain_a
n = 5
for (i in 1:length(webtrain_a)){
  if (webtrain_a[i] >= n){
    webtrain_a_n_5[i] = 1
  }
  else{
    webtrain_a_n_5[i] = webtrain_a_n_5[i]/5
  }
}

webtrain_a_matrix_5 = as.numeric(webtrain_a_n_5)
webtrain_a_matrix_5 = matrix(webtrain_a_matrix_5, nrow = 4151, ncol = 4151)
diag(webtrain_a_matrix_5) = 0
save(webtrain_a_matrix_5,file = "../output/webtrain_significance_weighting_n=5.Rdata")
```
### n=8
```{r}
webtrain_a_n_8 = webtrain_a
n = 8
for (i in 1:length(webtrain_a)){
  if (webtrain_a[i] >= n){
    webtrain_a_n_8[i] = 1
  }
  else{
    webtrain_a_n_8[i] = webtrain_a_n_5[i]/8
  }
}

webtrain_a_matrix_8 = as.numeric(webtrain_a_n_8)
webtrain_a_matrix_8 = matrix(webtrain_a_matrix_8, nrow = 4151, ncol = 4151)
diag(webtrain_a_matrix_8) = 0
save(webtrain_a_matrix_8,file = "../output/webtrain_significance_weighting_n=8.Rdata")
```
### n=9
```{r}
webtrain_a_n_9 = webtrain_a
n = 9
for (i in 1:length(webtrain_a)){
  if (webtrain_a[i] >= n){
    webtrain_a_n_9[i] = 1
  }
  else{
    webtrain_a_n_9[i] = webtrain_a_n_9[i]/9
  }
}

webtrain_a_matrix_9 = as.numeric(webtrain_a_n_9)
webtrain_a_matrix_9 = matrix(webtrain_a_matrix_9, nrow = 4151, ncol = 4151)
diag(webtrain_a_matrix_9) = 0
save(webtrain_a_matrix_9,file = "../output/webtrain_significance_weighting_n=9.Rdata")
```
### n=10
```{r}
webtrain_a_n_10 = webtrain_a
n = 10
for (i in 1:length(webtrain_a)){
  if (webtrain_a[i] >= n){
    webtrain_a_n_10[i] = 1
  }
  else{
    webtrain_a_n_10[i] = webtrain_a_n_10[i]/10
  }
}

webtrain_a_matrix_10 = as.numeric(webtrain_a_n_10)
webtrain_a_matrix_10 = matrix(webtrain_a_matrix_10, nrow = 4151, ncol = 4151)
diag(webtrain_a_matrix_10) = 0
save(webtrain_a_matrix_10,file = "../output/webtrain_significance_weighting_n=10.Rdata")
```
## The propotions of values that are not 1
```{r}
i=17230801
a1 = (i - sum(webtrain_a_n_5 == 1))/i
a2 = (i - sum(webtrain_a_n_8 == 1))/i
a3 = (i - sum(webtrain_a_n_9 == 1))/i
a4 = (i - sum(webtrain_a_n_10 == 1))/i
```
## Movie dataset
### We choose 9, 10, 11 and 12 as the number of co-rated items
### n = 9
```{r}
movie_train_a_9 = movie_train_a
n = 9
for(x in 1:dim(movie_train_a)[1]){
  for(y in 1:dim(movie_train_a)[1]){
    if(movie_train_a_9[x,y] >= n){
      movie_train_a_9[x,y] = 1
    }
    else{
      movie_train_a_9[x,y] = movie_train_a_9[x,y]/n
    }
  }
}

save(movie_train_a_9,file = "../output/movietrain_significance_weighting_n=9.Rdata")
```
### n = 10
```{r}
movie_train_a_10 = movie_train_a
n = 10
for(x in 1:dim(movie_train_a)[1]){
  for(y in 1:dim(movie_train_a)[1]){
    if(movie_train_a_10[x,y] >= n){
      movie_train_a_10[x,y] = 1
    }
    else{
      movie_train_a_10[x,y] = movie_train_a_10[x,y]/n
    }
  }
}

save(movie_train_a_10,file = "../output/movietrain_significance_weighting_n=10.Rdata")
```
### n = 11
```{r}
movie_train_a_11 = movie_train_a
n = 11
for(x in 1:dim(movie_train_a)[1]){
  for(y in 1:dim(movie_train_a)[1]){
    if(movie_train_a_11[x,y] >= n){
      movie_train_a_11[x,y] = 1
    }
    else{
      movie_train_a_11[x,y] = movie_train_a_11[x,y]/n
    }
  }
}

save(movie_train_a_11,file = "../output/movietrain_significance_weighting_n=11.Rdata")
```
### n = 12
```{r}
movie_train_a_12 = movie_train_a
n = 12
for(x in 1:dim(movie_train_a)[1]){
  for(y in 1:dim(movie_train_a)[1]){
    if(movie_train_a_12[x,y] >= n){
      movie_train_a_12[x,y] = 1
    }
    else{
      movie_train_a_12[x,y] = movie_train_a_12[x,y]/n
    }
  }
}

save(movie_train_a_12,file = "../output/movietrain_significance_weighting_n=12.Rdata")
```
### n = 14
```{r}
movie_train_a_14 = movie_train_a
n = 14
for(x in 1:dim(movie_train_a)[1]){
  for(y in 1:dim(movie_train_a)[1]){
    if(movie_train_a_14[x,y] >= n){
      movie_train_a_14[x,y] = 1
    }
    else{
      movie_train_a_14[x,y] = movie_train_a_14[x,y]/n
    }
  }
}

save(movie_train_a_14,file = "../output/movietrain_significance_weighting_n=14.Rdata")
```
### n = 16
```{r}
movie_train_a_16 = movie_train_a
n = 16
for(x in 1:dim(movie_train_a)[1]){
  for(y in 1:dim(movie_train_a)[1]){
    if(movie_train_a_16[x,y] >= n){
      movie_train_a_16[x,y] = 1
    }
    else{
      movie_train_a_16[x,y] = movie_train_a_16[x,y]/n
    }
  }
}

save(movie_train_a_16,file = "../output/movietrain_significance_weighting_n=16.Rdata")
```
## The propotions of values that are not 1
```{r}
i=25553025
a1 = (i - sum(movie_train_a_9 == 1))/i
a2 = (i - sum(movie_train_a_10 == 1))/i
a3 = (i - sum(movie_train_a_11 == 1))/i
a4 = (i - sum(movie_train_a_12 == 1))/i
a5 = (i - sum(movie_train_a_14 == 1))/i
a6 = (i - sum(movie_train_a_16 == 1))/i
```